---
title: "HMDA Census Tract Level Maps"
output: html_document
date: "2024-02-25"
---
The following are loan statistics from HMDA data reported at the census tract level.

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# https://stackoverflow.com/questions/61771543/assigning-colors-to-table-values-using-choroplethr

library(knitr)
library(choroplethr)
library(choroplethrMaps)
library(choroplethrZip)
delay_code_labels <- NULL
knit_hooks$set(delay = function(before, options, envir) {
    if (before) {
        delay_code_labels <<- append(delay_code_labels, options$label)
        return(NULL)  ## otherwise knitr will print delay_code_labels every time
    } else {
        
    }
})
opts_chunk$set(delay = TRUE, echo = FALSE, message = FALSE, warning = FALSE, 
              fig.align='left', fig.width = 8,fig.height=8,dpi=300)

```

```{r }
setwd('/home/shino/00 SF Loan Performance/00nano/89hmda')
data(state.regions)
#reference_map<-TRUE
reference_map<-FALSE
# when setting reference_map<-TRUE , see Google authorization notes below
# https://www.r-bloggers.com/2019/05/bug-when-creating-reference-maps-with-choroplethr/
library(ggmap)
library(mapproj)
register_google('AIzaSyBaOcyvB-2juG7eRhz79tQalWMtsAMFqh0')
#state_choropleth(df_pop_state,
#    num_colors = 1,
#    zoom = c("california", "oregon", "washington"),
#    reference_map = TRUE)
overwrite<-FALSE
state_list<-c('new york','new jersey') # ,'california','texas','florida')
outfile<-'cbsa.tsv'
cbsa_sqlfile<-'cbsa_v01.sql'
year<-2021
if(overwrite | !file.exists(outfile))
{
  system(paste0('cat ',cbsa_sqlfile, ' | cpp -I. -I.. -DYEAR=',year
                ,' | time mysql --login-path=shino loans | tee ',outfile))
}
cbsa<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
data(state.regions)
```

```{r}
map_tracts<-function(data,state,title,cbsa)
{
  library(ggplot2)
  library(RColorBrewer)
  breaks<-c(0,0.05,0.10,0.25,0.5,1)
  n<-length(breaks)-1
  palette<-"Reds"
  cap_state<-gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2",    # Uppercase with Base R
     state,
     perl = TRUE)
  map_title<-paste('Full State: ',cap_state,'\n',title)
  map_list<-list() # plots
  # do the state plot
  map_list<-append(map_list,list(
    tract_choropleth(data,state_name=state,title=map_title,num_colors=1
                     ,reference_map=reference_map) +
      scale_fill_stepsn(colors=RColorBrewer::brewer.pal(n,palette)
                                ,breaks=breaks,n.breaks=n
                                ,na.value=0,show.limits=TRUE)
    ))
  cbsa_list<-unique(cbsa[cbsa$state_name==cap_state,c('cbsa_code','cbsa_title')])
  # do select CBSA and Metro Division plots
  if(nrow(cbsa_list)>0)
  {
    for(i in 1:nrow(cbsa_list)) # CBSAs
    {
      cbsa_code<-cbsa_list[i,1]
      cbsa_name<-cbsa_list[i,2]
      md_list<-unique(cbsa[(cbsa$state_name==cap_state)&(cbsa$cbsa_code==cbsa_code)
                           &!is.na(cbsa$metrodiv_code)
                           ,c('metrodiv_code','metrodiv_title')])

      if(nrow(md_list)>0) # if there are Metro Divisions, loop through them
      {
        for(j in 1:nrow(md_list)) # plot each Metro Division seperately
        {
          md_code<-md_list[j,1]
          md_name<-md_list[j,2] 
          county_list<-unique(cbsa[cbsa$state_name==cap_state&cbsa$metrodiv_code==md_code
                                   ,c('state_county_code')])
          county_list<-county_list[!is.na(county_list)]
          county_names<-unique(cbsa[(cbsa$state_name==cap_state)&(cbsa$metrodiv_code==md_code)
                                    ,c('county')])
          county_names<-county_names[!is.na(county_names)]
          county_title<-gsub("County","",paste0(t(county_names),collapse=','))
          county_title<-gsub(" ,",",",county_title)
          map_title<-paste('State:',cap_state,'\n'
                           ,'CBSA:',cbsa_code,cbsa_name,'\n'
                           ,'Metro Division:',md_code,md_name,'\n'
                           ,'Counties:',county_title,'\n',title)
          map_list<-append(map_list,list(
            tract_choropleth(data,state_name=state,title=map_title,num_colors=1
                             ,county_zoom=county_list
                             ,reference_map=reference_map) +
              scale_fill_stepsn(colors=RColorBrewer::brewer.pal(n,palette)
                                ,breaks=breaks,n.breaks=n
                                ,na.value=0,show.limits=TRUE)
          ))
        }        
      }
      else # no Metro Divisions, plot counties in CBSA
      {
        county_list<-cbsa[cbsa$state_name==cap_state&cbsa$cbsa_code==cbsa_code,c('state_county_code')]
        county_names<-cbsa[cbsa$state_name==cap_state&cbsa$cbsa_code==cbsa_code,c('county')]
        county_title<-gsub("County","",paste0(t(county_names),collapse=','))
        county_title<-gsub(" ,",",",county_title)
        map_title<-paste('State:',cap_state,'\n'
                         ,'CBSA:',cbsa_code,cbsa_name,'\n'
                         ,'Counties:',county_title,'\n',title)
        map_list<-append(map_list,list(
          tract_choropleth(data,state_name=state,title=map_title,num_colors=1
                           ,county_zoom=county_list
                           ,reference_map=reference_map) + 
            scale_fill_stepsn(colors=RColorBrewer::brewer.pal(n,palette)
                              ,breaks=breaks,n.breaks=n
                              ,na.value=0,show.limits=TRUE)
          ))
      }
    }
  }
  return (map_list)
}

```

```{r hmda_strats}
script='run_sql_v22.sh'
sqlfile<-"hmda_strats_v06.sql"
param<-'CONV CONF PURCH YEARS=2019,2020,2021,2022,2023'
prefix<-'_0-'
suffix<-'.tsv'
outfile<-paste0(prefix,gsub(' ','-',param),suffix)
if(!file.exists(sqlfile))
{
  knitr::knit_exit()
}
if(overwrite | !file.exists(outfile))
{
  system(paste(script,sqlfile,param))
}
if(!file.exists(outfile))
{
  knitr::knit_exit()
}
hmda_data<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
```

## Second Homes

```{r second}

map_title<-paste("Second Homes ",param)
data<-hmda_data[,c('census_tract','second_home_pct')]
names(data)<-c("region","value")
maps<-list()
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,map_tracts(data,state_name,map_title,cbsa))
}
```

```{r}
for (map in maps)
{
  print(map)
}
```

<!-- ## Cashout Refi -->

<!-- ```{r cashout} -->

<!-- map_title<-paste("Cashout Refis",year) -->
<!-- maps<-list() -->
<!-- data<-hmda_data[,c('census_tract','cashout_pct')] -->
<!-- names(data)<-c("region","value") -->
<!-- ``` -->

<!-- ```{r, include=FALSE} -->
<!-- for (state_name in state_list) -->
<!-- { -->
<!--   maps<-append(maps,map_tracts(data,state_name,map_title,cbsa)) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for (map in maps) -->
<!-- { -->
<!--   print(map) -->
<!-- } -->
<!-- ``` -->

## Low Income (LMI) Loans

```{r lmi}
map_title<-paste("LMI Loans (<80% AMI) in",year)
maps<-list()
data<-hmda_data[,c('census_tract','lmi_pct')]
names(data)<-c("region","value")
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,map_tracts(data,state_name,map_title, cbsa))
}
```

```{r}
for (map in maps)
{
  print(map)
}
```

## Source Listing: `r knitr::current_input()`

```{r codeprint, echo = TRUE, eval = FALSE, ref.label = delay_code_labels, delay = FALSE}
```


## Query Listing: `r cbsa_sqlfile`

`r cbsa_sqlfile` queries the census delineation table and returns constituent counties comprising select CBSAs.
```sql
`r xfun::file_string(cbsa_sqlfile)`
```

## Query Listing: `r sqlfile`

`r sqlfile` queries hmda_lar and returns aggregate loan statistics grouped by census_tract for conventional loans.
```sql
`r xfun::file_string(sqlfile)`
```

## Environment
```{r}
sessionInfo()
```

