---
title: "HMDA Census Tract Level Maps"
output: html_document
date: "2024-02-25"
---
The following are loan statistics from HMDA data reported at the census tract level.

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# https://stackoverflow.com/questions/61771543/assigning-colors-to-table-values-using-choroplethr

library(knitr)
library(choroplethr)
library(choroplethrMaps)
library(choroplethrZip)
delay_code_labels <- NULL
knit_hooks$set(delay = function(before, options, envir) {
    if (before) {
        delay_code_labels <<- append(delay_code_labels, options$label)
        return(NULL)  ## otherwise knitr will print delay_code_labels every time
    } else {
        
    }
})
opts_chunk$set(delay = TRUE, echo = FALSE, message = FALSE, warning = FALSE, 
              fig.align='left', fig.width = 8,fig.height=8,dpi=300)

```


```{r}
map_tracts<-function(data,state,title,cbsa)
{
  library(ggplot2)
  library(RColorBrewer)
  cap_state<-gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2",    # Uppercase with Base R
     state,
     perl = TRUE)
  map_title<-paste('Full State: ',cap_state,'\n',title)
  map_list<-list() # plots
  # do the state plot
  map_list<-append(map_list,list(
    tract_choropleth(data,state_name=state,title=map_title,num_colors=6) 
    ))
  cbsa_list<-unique(cbsa[cbsa$state_name==cap_state,c('cbsa_code','cbsa_title')])
  # do select CBSA and Metro Division plots
  if(nrow(cbsa_list)>0)
  {
    for(i in 1:nrow(cbsa_list)) # CBSAs
    {
      cbsa_code<-cbsa_list[i,1]
      cbsa_name<-cbsa_list[i,2]
      md_list<-unique(cbsa[(cbsa$state_name==cap_state)&(cbsa$cbsa_code==cbsa_code)
                           &!is.na(cbsa$metrodiv_code)
                           ,c('metrodiv_code','metrodiv_title')])

      if(nrow(md_list)>0) # if there are Metro Divisions, loop through them
      {
        for(j in 1:nrow(md_list)) # plot each Metro Division seperately
        {
          md_code<-md_list[j,1]
          md_name<-md_list[j,2] 
          county_list<-unique(cbsa[cbsa$state_name==cap_state&cbsa$metrodiv_code==md_code
                                   ,c('state_county_code')])
          county_list<-county_list[!is.na(county_list)]
          county_names<-unique(cbsa[(cbsa$state_name==cap_state)&(cbsa$metrodiv_code==md_code)
                                    ,c('county')])
          county_names<-county_names[!is.na(county_names)]
          county_title<-gsub("County","",paste0(t(county_names),collapse=','))
          county_title<-gsub(" ,",",",county_title)
          map_title<-paste('State:',cap_state,'\n'
                           ,'CBSA:',cbsa_code,cbsa_name,'\n'
                           ,'Metro Division:',md_code,md_name,'\n'
                           ,'Counties:',county_title,'\n',title)
          map_list<-append(map_list,list(
            tract_choropleth(data,state_name=state,title=map_title,num_colors=6,county_zoom=county_list) 
          ))
        }        
      }
      else # no Metro Divisions, plot counties in CBSA
      {
        county_list<-cbsa[cbsa$state_name==cap_state&cbsa$cbsa_code==cbsa_code,c('state_county_code')]
        county_names<-cbsa[cbsa$state_name==cap_state&cbsa$cbsa_code==cbsa_code,c('county')]
        county_title<-gsub("County","",paste0(t(county_names),collapse=','))
        county_title<-gsub(" ,",",",county_title)
        map_title<-paste('State:',cap_state,'\n'
                         ,'CBSA:',cbsa_code,cbsa_name,'\n'
                         ,'Counties:',county_title,'\n',title)
        map_list<-append(map_list,list(
          tract_choropleth(data,state_name=state,title=map_title,num_colors=6,county_zoom=county_list) 
          ))
      }
    }
  }
  return (map_list)
}

  #  +
  #    scale_fill_manual(values=c("<1%" = "#ffffffff",  
  #                           "%1-5%" = "#64acbe", 
  #                           "5-10%" = "#c85a5a", 
  #                           ">10%" = "#574249",
  #                           "NA" = "#ffffffff"),
  #                  name="Second Homes")
```

```{r }
setwd('/home/shino/00 SF Loan Performance/00nano/89hmda')
data(state.regions)
overwrite<-TRUE
state_list<-c('new york','new jersey') # ,'california','texas','florida')
outfile<-'cbsa.tsv'
cbsa_sqlfile<-'cbsa_v01.sql'
if(overwrite | !file.exists(outfile))
{
  system(paste('cat',cbsa_sqlfile, ' | time mysql --login-path=shino loans | tee',outfile))
}
overwrite<-FALSE
cbsa<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
```

## Second Homes

```{r seconds}

outfile<-'second.tsv'
sqlfile<-"second_hmda_v01.sql"
if(overwrite | !file.exists(outfile))
{
  system(paste('cat',sqlfile, ' | time mysql --login-path=shino loans | tee',outfile))
}
data<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
data$orig<-data$value
#data$value <- cut (data$orig, breaks = c(0, 0.01, 0.05, 0.10, 1))

map_title<-"Second Homes in 2022"
maps<-list()
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,map_tracts(data,state_name,map_title,cbsa))
}
```

```{r}
for (map in maps)
{
  print(map)
}
```

## Cashout Refi

```{r cashout}

outfile<-'cashout_hmda_v01.tsv'
cash_sqlfile<-"cashout_hmda_v01.sql"
if(overwrite | !file.exists(outfile))
{
  system(paste('cat',cash_sqlfile, ' | time mysql --login-path=shino loans | tee',outfile))
}
data<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
data$orig<-data$value
#data$value <- cut (data$orig, breaks = c(0, 0.01, 0.05, 0.10, 1))

map_title<-"Cashout Refis in 2022"
maps<-list()
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,map_tracts(data,state_name,map_title,cbsa))
}
```

```{r}
for (map in maps)
{
  print(map)
}
```

## Low Income (LMI) Loans

```{r lmi}

outfile<-'lmi_hmda_v01.tsv'
sqlfile2<-"lmi_hmda_v01.sql"
if(overwrite | !file.exists(outfile))
{
  system(paste('cat',sqlfile2, ' | time mysql --login-path=shino loans | tee',outfile))
}
data<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
data$orig<-data$value
#data$value <- cut (data$orig, breaks = c(0, 0.01, 0.05, 0.10, 1))

map_title<-"LMI Loans (<80% AMI) in 2022"
maps<-list()
data(state.regions)
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,map_tracts(data,state_name,map_title, cbsa))
}
```


```{r}
for (map in maps)
{
  print(map)
}
```

## Source Listing: `r knitr::current_input()`

```{r codeprint, echo = TRUE, eval = FALSE, ref.label = delay_code_labels, delay = FALSE}
```


## Query Listing: `r cbsa_sqlfile`

`r cbsa_sqlfile` queries the census delineation table and returns constituent counties comprising select CBSAs.
```sql
`r xfun::file_string(cbsa_sqlfile)`
```

## Query Listing: `r sqlfile`

`r sqlfile` queries hmda_lar and returns second homes grouped by census_tract for conventional loans.
```sql
`r xfun::file_string(sqlfile)`
```
## Query Listing: `r cash_sqlfile`

`r cash_sqlfile` queries hmda_lar and returns cashout refis grouped by census_tract for conventional loans.
```sql
`r xfun::file_string(cash_sqlfile)`
```
## Query Listing: `r sqlfile2`

`r sqlfile2` queries hmda_lar and returns loans where qualifying income is less than 80% of area median income, grouped by census_tract for conventional loans.
```sql
`r xfun::file_string(sqlfile2)`
```
## Environment
```{r}
sessionInfo()
```

