---
title: "HMDA Census Tract Level Maps"
output: html_document
date: "2024-02-25"
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(choroplethr)
library(choroplethrMaps)
library(choroplethrZip)
delay_code_labels <- NULL
knit_hooks$set(delay = function(before, options, envir) {
    if (before) {
        delay_code_labels <<- append(delay_code_labels, options$label)
        return(NULL)  ## otherwise knitr will print delay_code_labels every time
    } else {
        
    }
})
opts_chunk$set(delay = TRUE, echo = FALSE, message = FALSE, warning = FALSE, 
               fig.width = 6, fig.asp = 0.618)

```


```{r}
map_tracts<-function(data,state,title)
{
  library(ggplot2)
  library(RColorBrewer)
  cap_state<-gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2",    # Uppercase with Base R
     state,
     perl = TRUE)
  map_title<-paste(cap_state,title)
  return (invisible(
    tract_choropleth(data,state_name=state,title=map_title,num_colors=6) 
      #+ scale_fill_manual(values = c("red", "grey", "seagreen3","yellow",na.value="white"))
      #,num_colors=5)
  #  +
  #    scale_fill_manual(values=c("<1%" = "#ffffffff",  
  #                           "%1-5%" = "#64acbe", 
  #                           "5-10%" = "#c85a5a", 
  #                           ">10%" = "#574249",
  #                           "NA" = "#ffffffff"),
  #                  name="Second Homes")
    ))
}
```

```{r }
setwd('/home/shino/00 SF Loan Performance/00nano/89hmda')
overwrite<-FALSE
state_list<-c('new york','new jersey') # ,'california','texas','florida')
#counties$"new york"<-c(36103,36059,36081,36047,36085,36061,36005,36119,36087)
#counties$"new jersey"<-c(34003,34017,34039)
```

```{r seconds}
# https://stackoverflow.com/questions/61771543/assigning-colors-to-table-values-using-choroplethr

outfile<-'second.tsv'
sqlfile<-"second_hmda_v01.sql"
if(overwrite | !file.exists(outfile))
{
  system(paste('cat',sqlfile, ' | time mysql --login-path=shino loans | tee',outfile))
}
data<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
data$orig<-data$value
#data$value <- cut (data$orig, breaks = c(0, 0.01, 0.05, 0.10, 1))

map_title<-"Conventional Mortgage Second Homes in 2022"
maps<-list()
data(state.regions)
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,list(map_tracts(data,state_name,map_title)))
}
```

## Second Homes

```{r}
for (map in maps)
{
  print(map)
}
```

```{r lmi}
# https://stackoverflow.com/questions/61771543/assigning-colors-to-table-values-using-choroplethr

outfile<-'lmi_hmda_v01.tsv'
sqlfile2<-"lmi_hmda_v01.sql"
if(overwrite | !file.exists(outfile))
{
  system(paste('cat',sqlfile2, ' | time mysql --login-path=shino loans | tee',outfile))
}
data<-read.csv(outfile,sep='\t') ## ,colClasses=c("character","numeric"))
data$orig<-data$value
#data$value <- cut (data$orig, breaks = c(0, 0.01, 0.05, 0.10, 1))

map_title<-"Conventional Mortgage LMI Loans (<80% AMI) in 2022"
maps<-list()
data(state.regions)
```

```{r, include=FALSE}
for (state_name in state_list)
{
  maps<-append(maps,list(map_tracts(data,state_name,map_title)))
}
```

## Low Income (LMI) Loans

```{r}
for (map in maps)
{
  print(map)
}
## Source File: `r knitr::current_input()`
```

## Source Listing: `r knitr::current_input()`

```{r codeprint, echo = TRUE, eval = FALSE, ref.label = delay_code_labels, delay = FALSE}
```

## Query Listing: `r sqlfile`

`r sqlfile` queries hmda_lar and returns second homes grouped by census_tract for conventional loans.
```sql
`r xfun::file_string(sqlfile)`
```
## Query Listing: `r sqlfile2`

`r sqlfile2` queries hmda_lar and returns loans where qualifying income is less than 80% of area median income, grouped by census_tract for conventional loans.
```sql
`r xfun::file_string(sqlfile2)`
```
## Environment
```{r}
sessionInfo()
```

